---
title: "Predictive Validity of the Unified Tertiary Matriculation Examination on University Grades"
subtitle: "A Validity Generalization Meta-Analysis"
toc: true
author: 
  - name: Matthew B. Jané
    affiliation:
      - name: University of Connecticut
  - name: James Uanhoro
    affiliation:
      - name: University of North Texas
  - name: Nduka Boika
    affiliation:
      - name: University of North Texas
  - name: Lucija Batinovic
    affiliation:
      - name: Linköping University
  - name: Blair T. Johnson
    affiliation:
      - name: University of Connecticut
  
description: "The Unified Tertiary Matriculation Examination (UTME) is a standardized test used for university admissions decisions in Nigeria. The examination is administered by the Joint Admissions and Matriculation Board (JAMB). The validity of the UTME in predicting academic performance has been investigated extensively in the literature, therefore this meta-analysis aims to synthesize the current validity evidence. This open-source web-page consists of the results, methodology, and data for a real-time updating meta-analytic database."
---
```{r,echo=FALSE,message=FALSE,results='asis'}
# load packages
library(psychmeta)
library(knitr)
library(ggplot2)
library(ggdist)
library(latex2exp)

# BLUE color pallette
text_color_blue      = '#326982ff'
panel_color_blue     = '#f6fafbff'
lightmain_color_blue = '#a4cdd9ff'
main_color_blue      = '#5fa6bcff'
darkmain_color_blue  = '#397689ff'
border_color_blue    = '#5fa6bcff'

# RED color pallette
text_color_red       = '#a62675ff'
panel_color_red      = '#fdf6faff'
lightmain_color_red  = '#eeb4d7ff'
main_color_red       = '#d74ea2ff'
darkmain_color_red   = '#bf2986ff'
border_color_red     = '#d74ea2ff'

# Plot theme BLUE
th_blue <- theme(aspect.ratio = 1,
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(color = text_color_blue),
        panel.background = element_rect(fill = panel_color_blue),
        panel.border = element_rect(fill = NA, color = border_color_blue,linewidth=1.2),
        axis.title = element_text(size=15, color = text_color_blue),
        axis.text.x = element_text(size=13, color = text_color_blue),
        axis.text.y = element_text(size=13, color = text_color_blue),
        axis.ticks = element_line(color = border_color_blue,linewidth=1)) 
  
# Plot theme RED
th_red <- theme(aspect.ratio = 1,
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(color = text_color_red),
        panel.background = element_rect(fill = panel_color_red),
        panel.border = element_rect(fill = NA, color = border_color_red,linewidth=1.2),
        axis.title = element_text(size=15, color = text_color_red),
        axis.text.x = element_text(size=13, color = text_color_red),
        axis.text.y = element_text(size=13, color = text_color_red),
        axis.ticks = element_line(color = border_color_red,linewidth=1)) 

```


# Current Results

## Study Selection

```{r,echo=FALSE,message=FALSE,results='asis',warnin}
library(PRISMAstatement)

sr <- read.csv('search_results.csv')

prisma(found = nrow(sr),
       found_other = 0,
       no_dupes = nrow(sr[sr$exclusion_reason!='duplicate',]), 
       screened = nrow(sr[sr$exclusion_reason!='duplicate',]), 
       screen_exclusions = 0, 
       full_text = nrow(sr[sr$exclusion_reason!='duplicate',]),
       full_text_exclusions = nrow(sr[sr$decision=='exclude',]), 
       qualitative = nrow(sr[sr$decision=='include',]), 
       quantitative = nrow(sr[sr$decision=='include',]),
       width = 800, height = 800)

```

## Imputing Missing Reliability

The average inter-correlation between each year-end GPA ($\bar{r}_{y_iy_j}$; see Methods) and the number of years that the students attend college, $\kappa$, can be used to estimate the reliability of final cumulative GPA (CGPA) during their time in college. For example, a CGPA calculated from all four years of college is more reliable than a CGPA computed from just their first year in college. Since inter-grade correlations are rarely reported in primary studies, we can take the $N$-weighted average of available inter-grade correlations to fill in missing values. We then use the number of years of university the study computes the CGPA from and use those values to estimate the reliability of CGPA in the given study.

```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
df <- read.csv('study_data.csv')

###################################################
# analyze data set from Popoola (2018) ############
###################################################
n <- c()
ave_ryiyj <- c()
for(i in 1:7){
  ryy_data <- readxl::read_xlsx('raw-data/popoola2018_gpa_only.xlsx',sheet = i)
  yiyj <- ryy_data[,c('First Year GPA','Second Year GPA','Third Year GPA','Fourth Year GPA','Fifth Year GPA')]
  cor_mat <- cor(yiyj)
  diag(cor_mat) <- NA
  ave_ryiyj[i] <- mean(cor_mat,na.rm=T)
  n[i] <- nrow(ryy_data)
}


###################################################
# analyze data set from Owolabi (2018) ############
###################################################

  ryy_data <- readxl::read_xlsx('raw-data/owolabi2018.xlsx',sheet = 1)
  yiyj <- ryy_data[,c('First Year GPA','Second Year GPA','Third Year GPA','Fourth Year GPA')]
  cor_mat <- cor(yiyj)
  diag(cor_mat) <- NA
  ave_ryiyj <- c(ave_ryiyj, mean(cor_mat,na.rm=T))
  n <- c(n,nrow(ryy_data))
  
  
reliability <- data.frame(
  n = c(n,df$n[!is.na(df$ave_ryiyj)]),
  ave_ryiyj = round(c(ave_ryiyj,df$ave_ryiyj[!is.na(df$ave_ryiyj)]),3)
)


# calculate n-weighted mean of inter-grade correlations
mean_ryiyj <- weighted.mean(x = reliability$ave_ryiyj,
                           w = reliability$n,
                           na.rm = T)


reliability_table <- rbind(reliability,c('WEIGHTED MEAN',sum(reliability$n),round(mean_ryiyj,3)))


```

#### Estimated Reliability

```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
#| hover: true

k = 1:5
rYY = c()
for(i in k){
  rYY[i] <- mean_ryiyj*i / (1+mean_ryiyj*(i-1))
} 

rYY_table <- data.frame(k,rYY)
names(rYY_table) <- c('# of Academic Years $(\\kappa)$','Estimated Reliability $(r_{YY\'})$')

kable(rYY_table,align = 'cc',digits = 3) 
```

## Full Meta-Analysis

The full meta-analysis of all studies is reported below with an uncorrected (i.e., bare bones) and corrected version.

#### Bare Bones Model (No Correction)

```{r,echo=FALSE,message=FALSE,results='asis'}
#| hover: true

rYY <- df$ave_ryiyj*df$k_grades / (1+df$ave_ryiyj*(df$k_grades-1))
df <- cbind(df,rYY)

admitted_all <- c(3708,29485,130865,257902,87988,10796,5628,0)
admitted_uni <- c(3708,29485,130865,257902,87988,0,0,0)
applicants <- c(5833,59494,320430,830866,407300,79837,80746,51874)
utme_score <- c(350,274.5,224.5,179.5,149.5,129.5,109.5,49.5)

SDapplicants <- sd(rep(utme_score,applicants))
SDadmitted_uni <- sd(rep(utme_score,admitted_uni))
SDadmitted_all <- sd(rep(utme_score,admitted_all))

df$sd_utme[is.na(df$sd_utme) & df$institution_type != "university"] <- SDadmitted_all
df$sd_utme[is.na(df$sd_utme) & df$institution_type == "university"] <- SDadmitted_uni
ux <- df$sd_utme / SDapplicants
df <- cbind(df,ux)

df$rxY[is.na(df$rxY)] <- composite_r_scalar(mean_rxy = df$ave_rxyj[is.na(df$rxY)], 
                                            k_vars_y = df$k_grades[is.na(df$rxY)], 
                                            mean_intercor_y = df$ave_ryiyj[is.na(df$rxY)])


mdl<- ma_r_ic(data = df,
              rxyi = rxY,
              n = n,
              ux = ux,
              ryy = rYY,
              wt_type = "PM",
              correct_bias = TRUE,
              correction_method = "uvirr_x",
              control = control_psychmeta(cred_level = .90)
              )



```

```{r,echo=FALSE,message=FALSE,results='asis'}
#| hover: true

summary_table <- as.data.frame(summary(mdl)$meta_tables$barebones)[, c('k','N','mean_r','CI_LL_95','CI_UL_95','sd_r','CR_LL_90','CR_UL_90')]

names(summary_table) <- c("$k$","$N$","   $\\bar{r}$   ", "$CI_{L95}$", "$CI_{U95}$","$SD_r$", "$CR_{L90}$", "$CR_{U90}$")

kable(summary_table,align = 'cccccccc',digits=3)
```

#### Artifact Correction Model

```{r,echo=FALSE,message=FALSE,results='asis'}
#| hover: true

summary_table <- as.data.frame(summary(mdl)$meta_tables$individual_correction$true_score)[, c('k','N','mean_rho','CI_LL_95','CI_UL_95','sd_rho','CR_LL_90','CR_UL_90')]

names(summary_table) <- c("$k$","$N$","$\\hat{\\bar{\\rho}}$", "$CI_{L95}$", "$CI_{U95}$","$SD_\\rho$", "$CR_{L90}$", "$CR_{U90}$")


kable(summary_table,align = 'cccccccc',digits=3)


```

## Moderators

#### Disciplines

College disciplines/majors are broken down into broad categories.

```{r,echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
#| hover: true


mdl<- ma_r_ic(data = df,
              rxyi = rxY,
              n = n,
              ux = ux,
              ryy = rYY,
              wt_type = "PM",
              correct_bias = TRUE,
              correction_method = "uvirr_x",
              moderators = discipline_category,
              control = control_psychmeta(cred_level = .90)
              )

summary_table <- as.data.frame(summary(mdl)$meta_tables$individual_correction$true_score)[, c('discipline_category','k','N','mean_rho','CI_LL_95','CI_UL_95','sd_rho','CR_LL_90','CR_UL_90')]
summary_table$CI_UL_95[summary_table$CI_UL_95 >1] = 1
summary_table$CR_UL_90[summary_table$CR_UL_90 >1] = 1
summary_table$CI_UL_95[summary_table$CI_UL_95 < -1] = -1
summary_table$CR_UL_90[summary_table$CR_UL_90 < -1] = -1

# names(summary_table) <- c("Discipline","$k$","$N$","$\\hat{\\bar{\\rho}}$", "$CI_{L95}$", "$CI_{U95}$","$SD_\\rho$", "$CR_{L90}$", "$CR_{U90}$")
# 
# kable(summary_table,align = 'lccccc',digits=3)


ggplot(data=summary_table, aes(x=mean_rho,y=discipline_category)) +
  geom_errorbarh(aes(xmax = CR_LL_90, xmin = CR_UL_90, 
                     height = 0),linewidth=1,color=main_color_red) + 
  geom_errorbarh(aes(xmax = CI_LL_95, xmin = CI_UL_95, 
                     height = 0),linewidth=2,color=darkmain_color_red) + 
  theme_minimal() + th_red +
  geom_point(shape=18,size=6,color=darkmain_color_red) +
  xlim(-.5,1) +
  geom_vline(xintercept = 0,linetype='dashed') +
  annotate(geom='text',
           x=summary_table$mean_rho,
           y=summary_table$discipline_category,
           label=round(summary_table$mean_rho,2),
           vjust=2.2, color=text_color_red) +
  theme(aspect.ratio=1) + 
  ylab('') +
  xlab(TeX('Predictive Validity (\\bar{\\rho})'))
```


```{r,echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
names(summary_table) <- c("Computer Based","$k$","$N$","$\\hat{\\bar{\\rho}}$", "$CI_{L95}$", "$CI_{U95}$","$SD_\\rho$", "$CR_{L90}$", "$CR_{U90}$")

kable(summary_table,align = 'lccccc',digits=3)
```


#### Type of Institution

Whether the institution is a university, polytechnic, or college of education.

```{r,echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
#| hover: true


mdl<- ma_r_ic(data = df,
              rxyi = rxY,
              n = n,
              ux = ux,
              ryy = rYY,
              wt_type = "PM",
              correct_bias = TRUE,
              correction_method = "uvirr_x",
              moderators = institution_type,
              control = control_psychmeta(cred_level = .90)
              )

summary_table <- as.data.frame(summary(mdl)$meta_tables$individual_correction$true_score)[, c('institution_type','k','N','mean_rho','CI_LL_95','CI_UL_95','sd_rho','CR_LL_90','CR_UL_90')]
summary_table$CI_UL_95[summary_table$CI_UL_95 >1] = 1
summary_table$CR_UL_90[summary_table$CR_UL_90 >1] = 1
summary_table$CI_UL_95[summary_table$CI_UL_95 < -1] = -1
summary_table$CR_UL_90[summary_table$CR_UL_90 < -1] = -1


ggplot(data=summary_table, aes(x=mean_rho,y=institution_type)) +
  geom_errorbarh(aes(xmax = CR_LL_90, xmin = CR_UL_90, height = 0),linewidth=1,color=main_color_red) + 
  geom_errorbarh(aes(xmax = CI_LL_95, xmin = CI_UL_95, height = 0),linewidth=2,color=darkmain_color_red) + 
  theme_minimal() + th_red +
  geom_point(shape=18,size=6,color=darkmain_color_red) +
  xlim(-.5,1) +
  geom_vline(xintercept = 0,linetype='dashed') +
  annotate(geom='text',
           x=summary_table$mean_rho,
           y=summary_table$institution_type,
           label=round(summary_table$mean_rho,2),
           vjust=3, color=text_color_red) +
  ylab('') +
  theme(aspect.ratio=1) + 
  xlab(TeX('Predictive Validity (\\bar{\\rho})'))
  
```


```{r,echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
names(summary_table) <- c("Institution Type","$k$","$N$","$\\hat{\\bar{\\rho}}$", "$CI_{L95}$", "$CI_{U95}$","$SD_\\rho$", "$CR_{L90}$", "$CR_{U90}$")

kable(summary_table,align = 'lccccc',digits=3)

```
